---
title: "wk7_prac_coding"
author: "chris"
date: "2021/11/24"
output: html_document
---

7.4.1 Setting up your Data
```{r library}
library(tidyverse)
library(tmap)
library(geojsonio)
library(plotly)
library(rgdal)
library(broom)
library(mapview)
library(crosstalk)
library(sf)
library(sp)
library(spdep)
library(car)
library(fs)
library(janitor)
```

```{r}
#download a zip file containing some boundaries we want to use

download.file("https://data.london.gov.uk/download/statistical-gis-boundary-files-london/9ba8c833-6370-4b11-abdc-314aa020d5e0/statistical-gis-boundaries-london.zip", 
              destfile="data/statistical-gis-boundaries-london.zip")
```

```{r}
#Get the zip file and extract it
listfiles<-dir_info(here::here("data")) %>%
  dplyr::filter(str_detect(path, ".zip")) %>%
  dplyr::select(path)%>%
  pull()%>%
  #print out the .gz file
  print()%>%
  as.character()%>%
  utils::unzip(exdir=here::here("data"))
```
```{r}
#look what is inside the zip

Londonwards<-dir_info(here::here("data", 
                                 "statistical-gis-boundaries-london", 
                                 "ESRI"))%>%
  #$ means exact match
  dplyr::filter(str_detect(path, 
                           "London_Ward_CityMerged.shp$"))%>%
  dplyr::select(path)%>%
  pull()%>%
  #read in the file in
  st_read()
```

```{r}
#check the data
qtm(Londonwards)
```

```{r}
#We can use readr to deal with the issues in this dataset - which are to do with text values being stored in columns containing numeric values

#read in some data - couple of things here. Read in specifying a load of likely 'n/a' values, also specify Latin1 as encoding as there is a pound sign (£) in one of the column headers - just to make things fun!

LondonWardProfiles <- read_csv("https://data.london.gov.uk/download/ward-profiles-and-atlas/772d2d64-e8c6-46cb-86f9-e52b4c7851bc/ward-profiles-excel-version.csv", 
                               na = c("", "NA", "n/a"), 
                               locale = locale(encoding = 'Latin1'), 
                               col_names = TRUE)
```

```{r}
#check all of the columns have been read in correctly
Datatypelist <- LondonWardProfiles %>% 
  clean_names()%>%
  summarise_all(class) %>%
  pivot_longer(everything(), 
               names_to="All_variables", 
               values_to="Variable_class")

Datatypelist
```
```{r}
#merge boundaries and data
LonWardProfiles <- Londonwards%>%
  left_join(.,
            LondonWardProfiles,
            by=c('GSS_CODE'='New code'))
```

```{r}
#map our dependent variable to see if the join has worked
tmap_mode('plot')
qtm(LonWardProfiles,
    fill="Average GCSE capped point scores - 2014",
    borders = NULL,
    fill.palette='Blues')

```
7.4.1.2 Additional Data
```{r}
#might be a good idea to see where the secondary schools are in London too
london_schools <- read_csv('https://data.london.gov.uk/download/london-schools-atlas/57046151-39a0-45d9-8dc0-27ea7fd02de8/all_schools_xy_2016.csv')

#from the coordinate values stored in the x and y columns, which look like they are latitude and longitude values, create a new points dataset
lon_school_sf <- st_as_sf(london_schools,
                          coords=c('x','y'),
                          crs=4326)

lon_sec_schools_sf <- lon_school_sf%>%
  clean_names()%>%
  filter(phase=='Secondary')

tmap_mode('plot')
qtm(lon_sec_schools_sf)
```

7.5 Analysing GCSE exam performance - testing a research hypothesis
To explore the factors that might influence GCSE exam performance in London, we are going to run a series of different regression models. A regression model is simply the expression of a linear relationship between our outcome variable (Average GCSE score in each Ward in London) and another variable or several variables that might explain this outcome.

7.5.1 Research Question and Hypothesis
```{r}
q <- qplot(x=`Unauthorised Absence in All Schools (%) - 2013`,
           y=`Average GCSE capped point scores - 2014`,
           data=LonWardProfiles)

#plot with a regression line - note, I've added some jitter here as the x-scale is rounded
q+stat_smooth(method = 'lm',se=FALSE,size=1)+
  geom_jitter()
```
7.5.3 Running a Regression Model in R
```{r}
#run the linear regression model and store its outputs in an object called model1
Regressiondata <- LonWardProfiles%>%
  clean_names()%>%
  dplyr::select(average_gcse_capped_point_scores_2014,
                unauthorised_absence_in_all_schools_percent_2013)

#now model
model1 <- Regressiondata%>%
  lm(average_gcse_capped_point_scores_2014~
       unauthorised_absence_in_all_schools_percent_2013,
     data = .)

summary(model1)
```

7.5.3.2 broom
```{r}
library(broom)
tidy(model1)
glance(model1)
```

```{r}
library(tidypredict)
Regressiondata %>%
  tidypredict_to_column(model1)
```

7.5.4 Bootstrap resampling
7.5.6 Assumption 1 - There is a linear relationship between the dependent and independent variables
```{r}
# use Janitor to clean up the names.

LonWardProfiles <- LonWardProfiles %>%
  clean_names()

#let's check the distribution of these variables first

ggplot(LonWardProfiles, aes(x=average_gcse_capped_point_scores_2014)) + 
  geom_histogram(aes(y = ..density..),
                 binwidth = 5) + 
  geom_density(colour="red", 
               size=1, 
               adjust=1)
```

```{r}
ggplot(LonWardProfiles, aes(x=unauthorised_absence_in_all_schools_percent_2013)) +
  geom_histogram(aes(y = ..density..),
                 binwidth = 0.1) + 
  geom_density(colour="red",
               size=1, 
               adjust=1)
```
```{r}
ggplot(LonWardProfiles, aes(x=median_house_price_2014)) + 
  geom_histogram()
```
```{r}
qplot(x = median_house_price_2014, 
      y = average_gcse_capped_point_scores_2014, 
      data=LonWardProfiles)
```

7.5.6.1 Transforming variables
```{r}
ggplot(LonWardProfiles, aes(x=log(median_house_price_2014))) + 
  geom_histogram()
```
```{r}
symbox(~median_house_price_2014, 
       LonWardProfiles, 
       na.rm=T,
       powers=seq(-3,3,by=.5))
```
```{r}
ggplot(LonWardProfiles, aes(x=(median_house_price_2014)^-1)) + 
  geom_histogram()
```

```{r}
qplot(x = (median_house_price_2014)^-1, 
      y = average_gcse_capped_point_scores_2014,
      data=LonWardProfiles)
```
```{r}
qplot(x = log(median_house_price_2014), 
      y = average_gcse_capped_point_scores_2014, 
      data=LonWardProfiles)
```

7.5.7 Assumption 2 - The residuals in your model should be normally distributed
```{r}
#save the residuals into your dataframe

model_data <- model1 %>%
  augment(., Regressiondata)

#plot residuals
model_data%>%
dplyr::select(.resid)%>%
  pull()%>%
  qplot()+ 
  geom_histogram() 
```
7.5.8 Assumption 3 - No Multicolinearity in the independent variables
```{r}
Regressiondata2<- LonWardProfiles%>%
  clean_names()%>%
  dplyr::select(average_gcse_capped_point_scores_2014,
         unauthorised_absence_in_all_schools_percent_2013,
         median_house_price_2014)

model2 <- lm(average_gcse_capped_point_scores_2014 ~ unauthorised_absence_in_all_schools_percent_2013 + 
               log(median_house_price_2014), data = Regressiondata2)

#show the summary of those outputs
tidy(model2)
```

```{r}
glance(model2)
```
```{r}
#and for future use, write the residuals out
model_data2 <- model2 %>%
  augment(., Regressiondata2)

# also add them to the shapelayer
LonWardProfiles <- LonWardProfiles %>%
  mutate(model2resids = residuals(model2))
```

```{r}
library(corrr)

Correlation <- LonWardProfiles %>%
  st_drop_geometry()%>%
  dplyr::select(average_gcse_capped_point_scores_2014,
         unauthorised_absence_in_all_schools_percent_2013,
         median_house_price_2014) %>%
  mutate(median_house_price_2014 =log(median_house_price_2014))%>%
    correlate() %>%
  # just focus on GCSE and house prices
  focus(-average_gcse_capped_point_scores_2014, mirror = TRUE) 


#visualise the correlation matrix
rplot(Correlation)
```
7.5.8.1 Variance Inflation Factor (VIF)
```{r}
vif(model2)
```

```{r}
position <- c(10:74)

Correlation_all<- LonWardProfiles %>%
  st_drop_geometry()%>%
  dplyr::select(position)%>%
    correlate()

rplot(Correlation_all)
```
7.5.9 Assumption 4 - Homoscedasticity
```{r}
#print some model diagnositcs. 
par(mfrow=c(2,2))    #plot to 2 by 2 array
plot(model2)
```
7.5.10 Assumption 5 - Independence of Errors
7.5.10.1 Standard Autocorrelation
```{r}
DW <- durbinWatsonTest(model2)
tidy(DW)
```

```{r}
#now plot the residuals
tmap_mode("view")
#qtm(LonWardProfiles, fill = "model1_resids")

tm_shape(LonWardProfiles) +
  tm_polygons("model2resids",
              palette = "PuBu") +
tm_shape(lon_sec_schools_sf) + tm_dots(col = 'type')
```

```{r}
#calculate the centroids of all Wards in London
coordsW <- LonWardProfiles%>%
  st_centroid()%>%
  st_geometry()

plot(coordsW)
```
```{r}
#Now we need to generate a spatial weights matrix 
#(remember from the lecture a couple of weeks ago). 
#We'll start with a simple binary matrix of queen's case neighbours

LWard_nb <- LonWardProfiles %>%
  poly2nb(., queen=T)

#or nearest neighbours
knn_wards <-coordsW %>%
  knearneigh(., k=4)

LWard_knn <- knn_wards %>%
  knn2nb()

#plot them
plot(LWard_nb, st_geometry(coordsW), col="red")
```
```{r}
plot(LWard_knn, st_geometry(coordsW), col="blue")
```
```{r}
#create a spatial weights matrix object from these weights

Lward.queens_weight <- LWard_nb %>%
  nb2listw(., style="W")

Lward.knn_4_weight <- LWard_knn %>%
  nb2listw(., style="W")
```

```{r}
Queen <- LonWardProfiles %>%
  st_drop_geometry()%>%
  dplyr::select(model2resids)%>%
  pull()%>%
  moran.test(., Lward.queens_weight)%>%
  tidy()
```

```{r}
Nearest_neighbour <- LonWardProfiles %>%
  st_drop_geometry()%>%
  dplyr::select(model2resids)%>%
  pull()%>%
  moran.test(., Lward.knn_4_weight)%>%
  tidy()

Queen
```
```{r}
Nearest_neighbour
```

7.6 Spatial Regression Models
7.6.1 Dealing with Spatially Autocorrelated Residuals - Spatial Lag and Spatial Error models
7.6.1.1 The Spatial Lag (lagged dependent variable) model
```{r}
#Original Model
model2 <- lm(average_gcse_capped_point_scores_2014 ~ unauthorised_absence_in_all_schools_percent_2013 + 
               log(median_house_price_2014), data = LonWardProfiles)

tidy(model2)
```

```{r}
library(spatialreg)

slag_dv_model2_queen <- lagsarlm(average_gcse_capped_point_scores_2014 ~ unauthorised_absence_in_all_schools_percent_2013 + 
               log(median_house_price_2014), 
               data = LonWardProfiles, 
               nb2listw(LWard_nb, style="C"), 
               method = "eigen")

#what do the outputs show?
tidy(slag_dv_model2_queen)
```

```{r}
#glance() gives model stats but this need something produced from a linear model
#here we have used lagsarlm()
glance(slag_dv_model2_queen)
```

```{r}
t<-summary(slag_dv_model2_queen)
```

```{r}
#run a spatially-lagged regression model
slag_dv_model2_knn4 <- lagsarlm(average_gcse_capped_point_scores_2014 ~ unauthorised_absence_in_all_schools_percent_2013 + 
               log(median_house_price_2014), 
               data = LonWardProfiles, 
               nb2listw(LWard_knn, 
                        style="C"), 
               method = "eigen")

#what do the outputs show?
tidy(slag_dv_model2_knn4)
```

```{r}
#write out the residuals

LonWardProfiles <- LonWardProfiles %>%
  mutate(slag_dv_model2_knn_resids = residuals(slag_dv_model2_knn4))

KNN4Moran <- LonWardProfiles %>%
  st_drop_geometry()%>%
  dplyr::select(slag_dv_model2_knn_resids)%>%
  pull()%>%
  moran.test(., Lward.knn_4_weight)%>%
  tidy()

KNN4Moran
```

7.6.1.2 The Spatial Error Model
```{r}
sem_model1 <- errorsarlm(average_gcse_capped_point_scores_2014 ~ unauthorised_absence_in_all_schools_percent_2013 + 
               log(median_house_price_2014), 
               data = LonWardProfiles,
               nb2listw(LWard_knn, style="C"), 
               method = "eigen")

tidy(sem_model1)
```

7.6.2 Key advice
```{r}
library(spdep)

Lward.queens_weight_ROW <- LWard_nb %>%
  nb2listw(., style="W")

lm.LMtests(model2, Lward.queens_weight_ROW, test = c("LMerr","LMlag","RLMerr","RLMlag","SARMA"))
```
7.6.4 More data
```{r}
extradata <- read_csv("https://www.dropbox.com/s/qay9q1jwpffxcqj/LondonAdditionalDataFixed.csv?raw=1")

#add the extra data too
LonWardProfiles <- LonWardProfiles%>%
  left_join(., 
            extradata, 
            by = c("gss_code" = "Wardcode"))%>%
  clean_names()

#print some of the column names
LonWardProfiles%>%
  names()%>%
  tail(., n=10)
```

7.6.5 Extending your regression model - Dummy Variables
```{r}
p <- ggplot(LonWardProfiles, 
            aes(x=unauth_absence_schools11, 
                y=average_gcse_capped_point_scores_2014))
p + geom_point(aes(colour = inner_outer)) 
```

```{r}
#first, let's make sure R is reading our InnerOuter variable as a factor
#see what it is at the moment...
isitfactor <- LonWardProfiles %>%
  dplyr::select(inner_outer)%>%
  summarise_all(class)

isitfactor
```

```{r}
# change to factor

LonWardProfiles<- LonWardProfiles %>%
  mutate(inner_outer=as.factor(inner_outer))

#now run the model
model3 <- lm(average_gcse_capped_point_scores_2014 ~ unauthorised_absence_in_all_schools_percent_2013 + 
               log(median_house_price_2014) + 
               inner_outer, 
             data = LonWardProfiles)
 
tidy(model3)
```

```{r}
contrasts(LonWardProfiles$inner_outer)
```

```{r}
LonWardProfiles <- LonWardProfiles %>%
  mutate(inner_outer = relevel(inner_outer, 
                               ref="Outer"))

model3 <- lm(average_gcse_capped_point_scores_2014 ~ unauthorised_absence_in_all_schools_percent_2013 + 
               log(median_house_price_2014) + 
               inner_outer, 
             data = LonWardProfiles)

tidy(model3)
```

7.7 Task 3 - Spatial Non-stationarity and Geographically Weighted Regression Models (GWR)
```{r}
#select some variables from the data file
myvars <- LonWardProfiles %>%
  dplyr::select(average_gcse_capped_point_scores_2014,
         unauthorised_absence_in_all_schools_percent_2013,
         median_house_price_2014,
         rate_of_job_seekers_allowance_jsa_claimants_2015,
         percent_with_level_4_qualifications_and_above_2011,
         inner_outer)

#check their correlations are OK
Correlation_myvars <- myvars %>%
  st_drop_geometry()%>%
  dplyr::select(-inner_outer)%>%
  correlate()

#run a final OLS model
model_final <- lm(average_gcse_capped_point_scores_2014 ~ unauthorised_absence_in_all_schools_percent_2013 +
```


```{r}
log(median_house_price_2014) +
```


```{r}
inner_outer + 
                    rate_of_job_seekers_allowance_jsa_claimants_2015 +
                    percent_with_level_4_qualifications_and_above_2011, 
                  data = myvars)

tidy(model_final)
```

```{r}
LonWardProfiles <- LonWardProfiles %>%
  mutate(model_final_res = residuals(model_final))

par(mfrow=c(2,2))
plot(model_final)
```

```{r}
qtm(LonWardProfiles, fill = "model_final_res")
```
```{r}
library(spgwr)

coordsW2 <- st_coordinates(coordsW)

LonWardProfiles2 <- cbind(LonWardProfiles,coordsW2)

GWRbandwidth <- gwr.sel(average_gcse_capped_point_scores_2014 ~ unauthorised_absence_in_all_schools_percent_2013 + 
                    log(median_house_price_2014) + 
                    inner_outer + 
                    rate_of_job_seekers_allowance_jsa_claimants_2015 +
                    percent_with_level_4_qualifications_and_above_2011, 
                  data = LonWardProfiles2, 
                        coords=cbind(LonWardProfiles2$X, LonWardProfiles2$Y),
                  adapt=T)
```
```{r}
GWRbandwidth
```

```{r}
#run the gwr model
gwr.model = gwr(average_gcse_capped_point_scores_2014 ~ unauthorised_absence_in_all_schools_percent_2013 + 
                    log(median_house_price_2014) + 
                    inner_outer + 
                    rate_of_job_seekers_allowance_jsa_claimants_2015 +
                    percent_with_level_4_qualifications_and_above_2011, 
                  data = LonWardProfiles2, 
                coords=cbind(LonWardProfiles2$X, LonWardProfiles2$Y), 
                adapt=GWRbandwidth,
                #matrix output
                hatmatrix=TRUE,
                #standard error
                se.fit=TRUE)

#print the results of the model
gwr.model
```
```{r}
results <- as.data.frame(gwr.model$SDF)
names(results)
```

```{r}
#attach coefficients to original SF


LonWardProfiles2 <- LonWardProfiles %>%
  mutate(coefUnauthAbs = results$unauthorised_absence_in_all_schools_percent_2013,
         coefHousePrice = results$log.median_house_price_2014.,
         coefJSA = rate_of_job_seekers_allowance_jsa_claimants_2015,
         coefLev4Qual = percent_with_level_4_qualifications_and_above_2011)
```

```{r}
tm_shape(LonWardProfiles2) +
  tm_polygons(col = "coefUnauthAbs", 
              palette = "RdBu", 
              alpha = 0.5)
```

```{r}
#run the significance test
sigTest = abs(gwr.model$SDF$"log.median_house_price_2014.")-2* gwr.model$SDF$"log.median_house_price_2014"

#sigTest=abs(gwr.model$SDF$'log.median_house_price_2014.')-2*gwr.model$SDF$'log.median_house_price_2014.'

#store significance results
LonWardProfiles2 <- LonWardProfiles2 %>%
  mutate(GWRUnauthSig = sigTest)
```

```{r}
tm_shape(LonWardProfiles2) +
  tm_polygons(col = "GWRUnauthSig", 
              palette = "RdYlBu")
```

